---
title: "Reverse Engineering Project"
author: "Alexandra alpert, Group 1"
date: "March 6 2024"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this notebook, we are reverse engineering the story, [STORY NAME](STORY%20URL)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
options(scipen=999)
library(tidyverse)
library(lubridate)
library(janitor)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data.

```{r}
murder_data <- read_csv("data/homicide-data.csv")


```

## Sentences to Engineer

In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

-   **Sentence text**: "City police made an arrest in 41 percent of homicides in 2014"
-   **Analysis summary**: When filtering for Baltimore in strictly 2014, the number given is 211. After this finding we used Str_detect to find which cases were closed, we did 86/211 which equals 40.7 which was the Washington Post reported.

```{r}
murder_data |>
  filter(str_detect(reported_date, "2014")) |>
  filter(str_detect(city, "Baltimore")) |>
  filter(str_detect(disposition, "Closed by arrest")) 

# Put code to reverse engineer sentence here
# Display results of code below this codeblock

```

### Sentence 2

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
murder_data |>
 filter(str_detect(reported_date, "2015") | str_detect(reported_date, "2016") | str_detect(reported_date, "2017"))|>
  filter(str_detect(city, "Baltimore")) |>
  filter (str_detect(disposition, "Closed by arrest"))

```

### Sentence 3

-   **Sentence text**: "Of 50 of the nation’s largest cities, Baltimore is one of 34 where police now make homicide arrests less often than in 2014, according to a Washington Post analysis. In Chicago, the homicide arrest rate has dropped 21 percentage points, in Boston it has dropped 12 points and in St. Louis it is down 9."
-   **Analysis summary**: We were able to confirm this finding by the Washington Post by comparing the 2014 data to the 2017 data in each city. For instance:
-   Chicago: 2014 (125/418)=29.9%, 2017 (56/654)=8.5% — 21% drop
-   St Louis 2014 (65/160)=40.6% , 2017 (64/202)=31.7% — 9% drop
-   Boston 2014 (29/55)=52.7%, 2017 (22/54)40.7% — 12% drop

```{r}
#160
murder_data |>
  filter(str_detect(reported_date, "2014")) |>
  filter(str_detect(city, "St. Louis")) 
#65
murder_data |>
  filter(str_detect(reported_date, "2014")) |>
  filter(str_detect(city, "St. Louis")) |>
  filter (str_detect(disposition, "Closed by arrest"))
#202
murder_data |>
  filter(str_detect(reported_date, "2017")) |>
  filter(str_detect(city, "St. Louis")) 
#64
murder_data |>
  filter(str_detect(reported_date, "2017")) |>
  filter(str_detect(city, "St. Louis")) |>
  filter (str_detect(disposition, "Closed by arrest"))

#418
murder_data |>
  filter(str_detect(reported_date, "2014")) |>
  filter(str_detect(city, "Chicago")) 
#125
murder_data |>
  filter(str_detect(reported_date, "2014")) |>
  filter(str_detect(city, "Chicago")) |>
  filter (str_detect(disposition, "Closed by arrest"))
#654
murder_data |>
  filter(str_detect(reported_date, "2017")) |>
  filter(str_detect(city, "Chicago")) 
#56
murder_data |>
  filter(str_detect(reported_date, "2017")) |>
  filter(str_detect(city, "Chicago")) |>
  filter (str_detect(disposition, "Closed by arrest"))


#55
murder_data |>
  filter(str_detect(reported_date, "2014")) |>
  filter(str_detect(city, "Boston")) 
#29
murder_data |>
  filter(str_detect(reported_date, "2014")) |>
  filter(str_detect(city, "Boston")) |>
  filter (str_detect(disposition, "Closed by arrest"))
#54
murder_data |>
  filter(str_detect(reported_date, "2017")) |>
  filter(str_detect(city, "Boston")) 
#22
murder_data |>
  filter(str_detect(reported_date, "2017")) |>
  filter(str_detect(city, "Boston")) |>
  filter (str_detect(disposition, "Closed by arrest"))

```

### Sentence 4

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-
