---
title: "lab_04"
author: "Derek Willis"
date: "2023-01-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About this lab

To complete this lab, you need to: \* run existing code as directed (look for **Task**). \* modify existing code as directed (look for **Task**). \* write code in empty codeblocks provided to answer questions included (look for **Q**). \* write out the answer in the form of a complete sentence in the space given (look for **A**). Read the questions carefully and attempt to do all of what they ask. Make a list if it helps.

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings

You'll need to load three packages for this: the tidyverse, lubridate and janitor.

**Task** load these three packages.

```{r}
# Turn off scientific notation
options(scipen=999)
library(tidyverse)
library(lubridate)
library(janitor)

```

For this lab, we want to investigate spending by Maryland state and local candidates during the 2022 election. For example, we could ask how much money have candidates and committees spent on advertising on Facebook? Which vendors received the most money for media expenses? We have the data, but it's messy - names of payees are spelled differently and capitalized differently - and we need to clean it up. We'll use a combination of RStudio and OpenRefine to do that.

The basic process here is to start in RStudio, export data so that we can use OpenRefine to clean it up, and then bring it back into RStudio to finish our analysis. The function to export data from RStudio is called `write_csv()` and it takes two arguments: the name of the dataframe and the path of the file to write it to, which is in quotes. Like `read_csv()`, but the variable name comes inside. Here's an example, in which you'd replace the dataframe and the path values.

```{r}
write_csv(dataframe, "folder/file.csv")
```

## Load Data

You'll need to load one data set, a CSV file of Maryland campaign expenditures located in the data folder called "maryland_expenses.csv"

**Task** Create a codeblock below, then read the data in and assign it to an appropriate variable name. You'll want to clean up the column names and make sure that any date columns are actually date datatypes.

```{r}
# Load the Maryland expenditures data table, clean the column names and convert columns with dates into actual dates.
md_expenses <-read_csv('data/maryland_expenses.csv') |> clean_names()
md_expenses <- md_expenses |> mutate(expenditure_date = mdy(expenditure_date))
```

## Answer questions

**Q1.** You've been assigned to write about Maryland political committees' use of Facebook advertising. There's a category column in the data that has values for different kinds of spending. One of them is "Media". Write code to isolate Media expenses where the name of the payee is not NA.

Then, using the `write_csv` function, make a CSV file of those media expenses in your data folder. Using OpenRefine, create a project using that CSV file, create a copy of the `payee_name` column called `payee_clean` and then standardize it, focusing your work on the payees that appear most often. Pay particular attention to variations of Facebook, bearing in mind that it changed its name in 2021: <https://about.fb.com/news/2021/10/facebook-company-is-now-meta/>).

Then, when you are done standardizing `payee_clean`, export it as a new CSV file to your data folder, *giving that file a different name* and read it back into RStudio as a new dataframe.

Then, using that `payee_clean` column, write code to find all of the spending where Facebook is the payee and calculate the total amount of spending for each committee and present the results showing the committee that spent the most first. Which committee is that?

**A1: top spender:** Jain Ashwani Friends of

```{r}
# One step at a time. Find the category column in the data that has values for different kinds of spending.

# One of them is "Media". Write code to isolate Media expenses where the name of the payee is not NA. They key here is using (!is.na(payee_name)) within a function

media_expenses <- md_expenses |> filter(expense_category =='Media') |> filter(!is.na(payee_name))

write_csv(media_expenses,'data/media_expenses.csv')

```

```{r}
# Export this table as a CSV
write_csv(media_expenses,'data/media_expenses.csv')



# Clean up work in OpenRefine

```

```{r}
# read the cleaned CSV 
media_expenses_clean <- read_csv('data/media_expenses_clean.csv')
# calculate top Facebook spenders

media_expenses_clean |>
filter(payee_clean == 'Facebook') |>
  group_by(committee_name) |>
  summarize(total_amount =sum(amount)) |>
  arrange(desc(total_amount))

```

**Q2.** Let's do some research on the top committees from question 1. Look up the top 10 spenders: if they are candidate committees, what offices were they running for? If they were not candidates, what issues do they seem to care about? Write some sentences answering those questions, then go the Facebook ad library ([https://www.facebook.com/ads/library/?active_status=all&ad_type=political_and_issue_ads&country=US&sort_data\\[direction\\]=desc&sort_data\\[mode\\]=relevancy_monthly_grouped&media_type=all](https://www.facebook.com/ads/library/?active_status=all&ad_type=political_and_issue_ads&country=US&sort_data%5Bdirection%5D=desc&sort_data%5Bmode%5D=relevancy_monthly_grouped&media_type=all){.uri}) and look for ads from the top spender. What do you notice about those ads?

1.  The top spender, "Jain Ashwani friends of" was running as a candidate for governer as a Democrat, interested in the middle east and aligns himself with most democratic beliefs
2.  Monique Anderson-Walker Franchot , a candidate for governer as a Democrat. All of his ads in the Facebook library were removed.
3.  Franchot Peter, a candidate for governer as a Democrat, also had all of his ads removed and had minimal information about him on the internet
4.  Republican candidate for Harford, the ads are about small businesses and preserving farms
5.  MD Governer candidate, could not find any ads with the exact name on the Facebook library
6.  Working families national PAC, no ads and confused about if this is a candidate commitee
7.  Maryland Governor candidates, John King and Michelle Siri. King was formally a Secretary and has multiple ads and endorsements.
8.  Democratic Governor candidate for Maryland in 2021. Has numerous ads speaking of his personal life and his personal beliefs.
9.  Ran for county council district five, had some ads asking for support.
10. union, which paid for ads hoping to convince people to vote against "Question A."

**A2. DO AT HOME**

**Q3.** Given the list of top spenders in A1 and the primary results for governor (<https://elections.maryland.gov/elections/2022/primary_results/gen_results_2022_1.html>), who is missing from that A1 list? Did those campaigns run Facebook ads, according to Facebook's ad library? Who are the biggest recipients of media expenses by the Democratic primary winner for governor (HINT: that person has two committees)?

**Answer**: Given the primary results, Wes Moore is missing from the list. After getting rid of the filter by facebook, Moore has the highest expenses. He did not use facebook ads, he used mainly TV and some online advertising.

**A3. DO AT HOME**

```{r}
media_expenses_clean |>
  group_by(committee_name) |>
  summarize(total_amount =sum(amount)) |>
  arrange(desc(total_amount))

```

**Q4.** The media expense data you cleaned also has a standardized purpose column. Write code to generate a list of the committees with the total amount of spending on online advertising. How does this list compare to the one from Q1 about top Facebook spenders? Of the top 5 committees who spent the most on Facebook ads from Q1, are there any that appear to have spent *all* of their online advertising dollars on Facebook?

**Answer:** 4a. The top facebook spenders are not the top spenders of overall online advertising. They also spent more than facebook ads did. 4b. After looking at the spreadsheet none of the top 5 spenders in Online Advertising used facebook. Jain Ashwani Friends of spent all of her advertising dollars on facebook.

**A4.**

```{r}
media_expenses_clean |>
  filter(expense_purpose == 'Online Advertising') |>
  group_by(committee_name, expense_purpose, payee_clean) |>
  summarize(total_amount =sum(amount)) |>
  arrange(desc(total_amount))


```

**Q5** You've been asked to summarize your findings about spending on Facebook by Maryland political candidates. How much do we know and what don't we know? What's the best story idea you've found here?

**A5:** candidates either put a lot of focus on facebook spending or the complete opposite. A perfect example of this is Wes Moore, who won and became governor who did not spend a dime on facebook advertising. I find this an interesting story - the coloration between facebook spending and election success.
